---
title: "Lab 7 Hotspots"
author: "Sara Orofino"
date: "5/22/2019"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Packages:
```{r, message=FALSE}
# Load packages:

library(raster)
library(rgdal)
library(rasterVis)
library(maps)
library(rgeos)
library(dplyr)
library(RColorBrewer)


```

###Research Questions  

 1. What is the spatial distribution of global fishing and land-based threats in the CA Current?  
 2. Where were MPAs placed? Which threats do they address?  
 
###Part 1

Set color scheme and import raster data:  

```{r}
# Rainbow colors 

cols = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255))

# set smaller margins for plotting
par(mar=c(2,2,1,1))

# Import raster data

all_threats <- raster("/Users/saraorofino/Documents/GitHub/CPLab7_Hotspots/Threats_data/full_modelnv.tif")

```

Visualize:  

```{r}

# First plot just the CA current data:

plot(all_threats, col=cols) 

# Add a basemap using the map package:
# ext=extent() is giving you the window you're viewing in this case lat/long
# add=T tells R to add it to the existing plot, fill and col are telling R to fill it in and with what color  

plot(all_threats, ext=extent(-130,-110,24,50), col=cols)
map('world', fill=T, add=T, col='gray')


```

Extent and Zoom:  

In the console if you type the raster layer you see:  
 - Class: what type of data it is (rasterlayer)  
 - Dimensions: Gives #rows, #columns, and total number of cells (rows * columns)
 - Resolution: resolution of each of the pixels (in the coorrdinate reference line - in this case lat/long)  
 - Extent: minimum and maximum of data set in units of the data (in this case lat/long)   
 - Coord. Ref: projection and datum, where it starts        
 - Data Source: where the file is located   
 - Names: what you named it when you imported it     
 
 Visualize a different extent, or subset, of the data:  
 
```{r}

# Plot a good extent for Santa Barbara Channel 
# Add a title using main

plot(all_threats, col=cols, ext=extent(-121,-117,32,35), main = "Cumulative Threats")

# Can use the zoom() function to have an interactive where you can zoom in by clicking
# NOTE - this doesn't work with pdf markdowns 


```
 
Look at the distribution of the data with a histogram:  

```{r, warning=FALSE}
# Histogram to see the shape of data distribution

hist(all_threats, main = "Cumulative Threats Frequency")

# Note there are like 16,000 cells in this data so the histogram only shows a subset of about 1000

# cellStats() allows us to run basic statistics

cellStats(all_threats, mean) #mean = 14.00092
cellStats(all_threats, sd) #sd = 2.9589
```

###Part 2 Raster Calculations   

Task: Overlay the top 20% of cumulative threats with the top 20% of species richness to find threat hotspots.  
This requires four steps:  
 1. Import the data (already done for threats)  
 2. Resample the data to the same resolution  
 3. Reclassify the data to find the top 20%  
 4. Overlay the data and find hotspots  
 
**Step 1. Import the data**    

```{r}

#Pull species data file:   
all_spp <- raster("/Users/saraorofino/Documents/GitHub/CPLab7_Hotspots/ca_curr_sp_rich.tif")

# Visualize
plot(all_spp, col=cols)

# Notice the difference in resolution - larger resolution looks more pixelated than the threats data 
```

Note: calling the file in the console shows the extent and resolution are different than the threats layer - makes raster calculations impossible   

**Step 2. Resample the data**  

```{r}

# Use crop() to change the extent - threats layer has the larger extent so we want to crop it to the same extent as the spp layer - crop to lowest common denominator   

threats_crop <- crop(all_threats, all_spp)


# Use resample() to make the resolutions the same - resample the spp layer based on the cropped threats layer 
# method = 'ngb' specifies we want to use a nearest neighbor algorithm to resample (instead of interpolation)
# Note: default for method is bilinear, need to specify for other methods  
# progress = 'text' will show the progress in the console of longer running functions --> great tool 

spp_res <- resample(all_spp, threats_crop, method = 'ngb', progress = 'text')

# Use stack() to create a RasterStack object - checks that the layers will line up decently 
spp_threat_stack <- stack(threats_crop, spp_res)
plot(spp_threat_stack, col=cols)

```

**Step 3. Reclassify**  

We want to find the top 20% of cells in each layer and then combine the two layers to produce hotspots. We can do this using the reclassify() function in the raster package.  

Function reclassify()  
rcl argument - matrix for reclassification; requires 3 columns    
 - First two columns are the "from" and "to" of the input values   
 - Third column is "becomes"; the new value for your range    
include.lowest argument - should the lowest value in the rcl be included?  
 - Default is false   
right arguement - should intervals be closed on the right (and open on the left) or vice versa  
 - Default is true   
 - right = NA will leave both intervals open  

Reclassify Species Richness:   
```{r}
# First look at the data in a histogram:  
hist(spp_res, main="Species Raster Values")
# Notice there are a lot of values at zero which should be NA 

# Reassign zero values to NA using reclassify 
spp_res <- reclassify(spp_res, rcl = c(-Inf, 0, NA))

# Histogram to check that the reclassification worked 
hist(spp_res, main="Species Raster Values, Zeros Removed")

# Find the top 20% and assign those cell values of 1 and all other non-NA values of zero using quantile()
# Find the value of the 80th percentile 
spp_cutoff <- quantile(spp_res, 0.80)

# Find the maximum value
spp_maxVal <- cellStats(spp_res, max)

# Reclassification Matrix (rcl) - everything from negative infinity to the cutoff 80th percentile is 0 and everything above cutoff to the maximum value is 1
rcl_mat <- c(-Inf, spp_cutoff, 0,
             spp_cutoff, spp_maxVal,1)

# Reclassify the species layer 
spp_binary <- reclassify(spp_res, rcl = rcl_mat)

# Visualize this to see if it worked:
binary_cols <- c("white", "firebrick")

plot(spp_binary, col=binary_cols, legend=F, main="Top 20% of Species Richness")
map('world', fill=T, add=T,col='gray')
```
 
Reclassify Threats:  
```{r}

# First look at the data in a histogram:  
hist(threats_crop, main="Threats Raster Values")
# Looks fine doesn't need to be reclassified to zero 

# Find the top 20% and assign those cell values of 1 and all other non-NA values of zero using quantile()
# Find the value of the 80th percentile 
threat_cutoff <- quantile(threats_crop, 0.80)

# Find the maximum value
threat_maxVal <- cellStats(threats_crop, max)

# Reclassification Matrix (rcl) - everything from negative infinity to the cutoff 80th percentile is 0 and everything above cutoff to the maximum value is 1
rcl_mat_threat <- c(-Inf, threat_cutoff, 0,
             threat_cutoff, threat_maxVal,1)

# Reclassify the species layer 
threat_binary <- reclassify(threats_crop, rcl = rcl_mat_threat)

# Visualize it to check if it worked:   
plot(threat_binary, col=binary_cols, legend=F, main="Top 20% of Cumulative Threats")
map('world', fill=T, add=T,col='gray')


```

**Step 4. Overlay for Hotspots**    

```{r}

# Overlay to find the hotspots - function tells R how you want to add them up in this case just x+y (adding zeros and ones)

hotspots <- overlay(spp_binary, threat_binary, fun=function(x,y){x+y})

# Color the breakpoints - need three colors now (cell values of 0,1,2)
#brks_hotspots <- seq(0,3,length.out = 4)
hotspot_cols <- c("white", "lightblue", "firebrick")

# Plot the hotspots 
plot(hotspots, col=hotspot_cols, legend=F, main="Hotspots")
map('world', fill=T, add=T, col='gray80')

# Plot just hotspots in the SB Channel by changing the extent 
plot(hotspots, col=hotspot_cols, ext=extent(-121,-117,32,35), main="Hotspots in SB Channel", legend=F)
map('world', fill=T, add=T, col='gray80')
```




